Efficient and Scalable Bipartite Matching with Fast Beta Linkage (fabl)

### 2023-05-02 ba_resubmit/response_to_reviewers2.tex

\documentclass[letterpaper, parskip]{scrartcl}

\usepackage{amsmath,amssymb,bm}
\usepackage{url}
\usepackage[hidelinks]{hyperref}
\usepackage[margin=1.4in]{geometry}
\usepackage{natbib}
\usepackage{xcolor}
\usepackage{tcolorbox}            % provides shaded box for reviewer comments
\usepackage{enumitem}             % use more compact lists
%\usepackage[parfill]{parskip}     % use skips between paragraphs and no indents
\usepackage{bibentry}             % provides \nobibliography command
\usepackage{xspace}

% Make itemize and enumerate environments more compact
\setlist[itemize]{noitemsep, topsep=0pt, parsep=0pt}
\setlist[enumerate]{noitemsep, topsep=0pt, parsep=0pt}

% -------- Define point raised command --------- %

\definecolor{darkpurple}{HTML}{500050}
\tcbuselibrary{breakable}       % allow tcolorbox to break across pages
\tcbuselibrary{skins}
\tcbset{%
	enhanced,%
	frame hidden,%
	sharp corners,%
	breakable,%
	parbox=false,%
	colback=lightgray!30,%
	enlarge left by=6pt,%
	width=\linewidth-6pt,%
	colback=gray!0,
	left=4pt,%
	right=0pt,%
	top=0pt,%
	bottom=0pt,%
	borderline west={2pt}{0pt}{lightgray},%
	%coltext=darkpurple,%
	before skip=14pt plus 2pt,%
	after skip=14pt plus 2pt}

%\newcommand{\pointRaised}[1]{%
%	\begin{tcolorbox}
%		\itshape #1
%	\end{tcolorbox}
%}
%
%\newcounter{responsectr}[section]     % counter for response resets at each section
%\newcommand{\reply}[2]{%
%	\refstepcounter{responsectr}%
%	\textbf{#1.\theresponsectr:} #2
%}

\newcommand{\pointRaised}[2]{%
	\textbf{#1.\theresponsectr:} #2
}

\newcounter{responsectr}[section]     % counter for response resets at each section
\newcommand{\reply}[1]{%
	\refstepcounter{responsectr}%
		\begin{tcolorbox}
			\itshape #1
		\end{tcolorbox}
}

\newcommand{\todo}{\textcolor{red}{[TODO]}\xspace}

\newcommand{\ncomment}[1]{{\color{teal} NM: #1}}

\begin{document}

	% ****************** TITLE ****************************************

	\title{Response to Reviewer Comments}
	\maketitle

Thank you for the thoughtful feedback. We agree that much of the notation iscumbersome, and we in particular thank the reviewers for seeing the merit in our methods despite those difficulties.  We have made several broad changes to address the Associate Editor's overall critiques, and have addressed each specific comment from the reviewer.

\section{AE Comments}

\pointRaised{AE}{I must add that reading the paper still feels like a chore, in part because the notation is so heavy, which may be hard to avoid, but also, in part, because some wording choices make certain sentences hard to parse, which can certainly be improved.}

\reply{We agree that our notation is at times cumbersome. We have made several efforts to simplify notation in this revision.

\begin{itemize}
	\item Instead of referring to files as $\mathbf{X}_1$ and $\mathbf{X}_2$, we are now using $A$ and $B$. We find this generally more readable, and particularly simplifies the notation in Section 4.2. 
	\item At the beginning of Section 4.3, we have rewritten the likelihood function in terms of the agreement patterns and the summary statistics discussed in Section 4.1. This is the text added:

''To ground the discussion of posterior inference, we first re-express the likelihood function from (6a) in terms of the agreement patterns and summary statistics in Table 2. Let $m(p) =  p(\gamma_{ij} = h_p|Z_j = i)$ be probability that records $A_i$ and $B_j$ form agreement pattern $p$ given that they are a match. Similarly define $u(p)$. These probabilities are simply products of relevant $\bm{m}$ and $\bm{u}$ parameters. Viewed through the perspective of patterns, the likelihood is
\begin{align}
	\mathcal{L}(\bm{Z}, \bm{m}, \bm{u} \mid \tilde{\bm{\Gamma}}) = \prod_{j=1}^{n_B}\prod_{p=1}^P \prod_{i \in r_{p_j}} m(p)^{I(Z_j = i)}u(p)^{1 - I(Z_j = i)}. \label{likelihood_efficient}"
\end{align}

	This likelihood is then used throughout the discussion of efficient posterior inference. We hope this expression clarifies the meaning of "preservation of weight (Comment 5), and strengthens the argument for $\tilde{\Gamma} = \{\mathcal{P}, \mathcal{R}, \mathcal{N} \}$ being sufficient statistics for $\bm{m}$, $\bm{u}$, and $\bm{Z}$ (Comment 4).

	\item To denote when the $(i,j)$ record pair exhibits agreement pattern $p$, we have removed the clunky $(i, j) \in h_p$ notation for $\gamma_{ij} = h_p$, which we feel is more direct and intuitive.

	\item Section 4.1 introduces the sufficient statistic $N_{j_p}$. After careful review, we realized that by changing the order of subscripts and instead using $N_{p_j}$, we can replace the sum of counts across all records (previously denoted $H_p$), with the more intuitive $N_p$. Thus, we have removed all of the ''$H$" quantities throughout Section 4. We believe this substantially improves readability.  
\end{itemize}
}

\pointRaised{AE}{Line 80 on p. 3 says: “For ease of readability, we follow the convention established by Sadinle (2017) and say “record $i \in X_1$” rather than the more compact $x_{1i}$.” Taking this at face value implies that, in any sentence, one could swap “record $i \in X_1$” for “$x_{1i}$,” which clearly is not true.  So, as it stands, the sentence does little to improve ease of readability.  In fact, if the point is that “$x_{1i}$” will never be used again, why is the notation introduced in the first place?}
\reply{We added this sentence in an attempt to address a previous reviewer comment pointing out a conflation between the record itself and the index directing towards that record. We did in fact intend for “record $i \in X1$” to be mean precisely $x_1i$. However, we see now that introducing $x_{1i}$ has introduced new confusion. With the simplified reference to the two datasets as $A$ and $B$, that material now reads:

	``Consider a data files $A$ and $B$, consisting of records $A_i$ and $B_j$ respectively, where $i \in \{1, \ldots n_A\}$ and $j \in \{1, \ldots n_B\}$".
}

\setcounter{responsectr}{0}
\section{Reviewer Comments}

	\pointRaised{R}{%
	For the missing data treatment now included just before Section 2.1, is the assumption truly
	missing at random, or missing completely at random? Or, does this distinction not matter
	because of the independence assumed across elements of the comparison vector?}

	\reply{%
\textcolor{red}{I used MAR because that is what Sadinle wrote in his paper. Please review, and confirm whether its MCAR or not.}

The assumption is missing completely at random. You are correct that this is a result of the independence assumed across elements of the comparison vector. This has been clarified.
	}

	\pointRaised{R}{%
	Appendix 8.2: I appreciate the streamlining of the discussion of the proposed algorithms and
	think it was a good choice to move the full derivation to an appendix. However, I do not follow
	the re-expression of the pmf for $\Gamma_{.j}$. First, the opening square brackets are still misplaced
	throughout. In the second line of this derivation, the authors divide the expression by a product
	of the element-wise conditional probabilities of a match in the comparison vector (u), raised by
	an indicator that the element of the comparison vector equals a particular value. Perhaps I
	misunderstand, but I believe this product is not constant in $\Gamma_{.j}$ or $u$, and so the total expression
	is not proportional to the line above, as interpreted as a pmf. I think that the end effect is that
	the last line of the newly expressed pmf is missing a factor of $\prod_i \prod_f \prod_l u_{fl}^{I(\gamma_{ij}^f = l) I_{obs}(\gamma_{ij}^f)}$ regardless of the value of $z_j$.}

\reply{%
	The square brackets have been corrected, and we apologize for the oversight.

	We have provided an alternative derivation of the full conditional, separately handling the cases for $Z_j \leq n_A$ and for $Z_j = n_A + j$. We provide it here for your reference:

	We consider the likelihood in two cases. When $B_j$ does not link to any record in $A$, the contribution to the likelihood is simply a product of $u$ parameters, which we will call $c_j$:
	\begin{align}
		p(\Gamma_{.j}| \bm{m}, \bm{u}, Z_j = n_A + j) = \prod_{i=1}^{n_A}\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)} = c_j.
	\end{align}
	When $Z_j = i$ for some $i \leq n_A$, we have
	\begin{align}
		p(\Gamma_{.j}| \bm{m}, \bm{u}, Z_j = i) =\prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)} \times \prod_{i' != i}\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{I(\gamma_{i'j}^f = l)I_{obs}(\gamma_{i'j}^f)}.
	\end{align}
	We multiply and divide by the $u$ parameters for the matching record pair to obtain
	\begin{align}
		p(\Gamma_{.j}| \bm{m}, \bm{u}, Z_j = i) &=\prod_{f=1}^{F}\prod_{l=1}^{L_f} \left(\frac{m_{fl}}{u_{fl}}\right)^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)} \times \prod_{i' = 1}^{n_A}\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{I(\gamma_{i'j}^f = l)I_{obs}(\gamma_{i'j}^f)} \\
		&= w_{ij} \times c_j.
	\end{align}
	Combining these results, we have 
	\begin{align}
		p(\Gamma_{.j}| \bm{m}, \bm{u}, Z_j = z_j) = \begin{cases} 
		w_{z_j, j}  & z_j \leq n_A; \\
		1 &  z_j  = n_A + j, \\
	\end{cases}
	\end{align}
}

%	{\color{red}{I don't really understand his confusion. Why wouldn't the product of $u$ parameters be constant in $u$ and the data? Perhaps if I use $\mathcal{L}(Z_j| m, y, \Gamma_{.j})$, it might be clearer?}}}

%	In the second line of the derivation, we divide by the $u$ parameters for all of the records in $A$ (note the product outside of the fraction). This quantity does not depend on $Z_j$ and is constant in $u$ and $\Gamma_{.j}$.  This is easiest to see through the likelihood expressed at the level of the agreement patterns. We are also notating this likelihood as $\mathcal{L}(Z_j | m, u, \Gamma_{.j})$ to make it clear what is considered constant.
%	
%	\begin{align*}
%	 \mathcal{L}(Z_j | m, u, \Gamma_{.j})  &= \prod_{p=1}^P \prod_{i \in r_{p_j}} m(p)^{I(Z_j = i)}u(p)^{1 - I(Z_j = i)}\\
%	 &= \prod_{p=1}^P u(p) ^ {N_{p_j}} \prod_{i \in r_{p_j}} m(p)^{I(Z_j = i)}u(p)^{- I(Z_j = i)}\\
%	 &= \prod_{p=1}^P u(p) ^ {N_{p_j}} \prod_{i \in r_{p_j}} w_p^{I(Z_j = i)}\\
%	\end{align*} 
%where $w_p = \frac{m(p)}{u(p)}$. The term outside of the inner product does not depend on $Z_j$. Thus we have
%\begin{align*}
%	\mathcal{L}(Z_j | m, u, \Gamma_{.j})   \propto
%	\begin{cases} 
%		w_p,  & z_j \leq n_1; \\
%		1, &  z_j  = n_1 + j; \\
%	\end{cases} \label{eqn:z}\\
%\end{align*}
%
%REVISIT

	\pointRaised{R}{%
Appendix 8.2: Thank you for including the details on integrating out pi from the full
conditionals. They surprised me. I had assumed that the authors had integrated out pi in the
prior for Z, as this is what Sadinle (2017) had done to form the “beta prior for bipartite
matchings”. Does the alternate approach presented in this paper provide a different algorithm
than directly integrating the prior distribution? Is the presented alternative approach justified?}

\reply{%
 {\color{red}{I really don't know how to answer this. I don't know how I would ''directly integrate" $\pi$, and I don't know if what I do is ''justified." Its kind of an intellectual relic to how I was originally thinking about the problem, and it reveals a quantity similar to what Sadinle was working with.

 It is theoretically more straightforward to just not integrate out $\pi$, and use samples from $\pi$ in Gibbs sampler like I do with all other parameters. Would that be too large a change at this stage?

 Without integrating out $\pi$, the full conditional is given on page 26 of the Appendix.}}
}

\pointRaised{R}{%
	I do not understand the statement just below Equation (10): “When j has no match in $X_1$, we write $(n_1 +j, j) \in h_{P+1}$” My understanding of these patterns is that they are
	based on observed comparison vectors without consideration of Z (matches). In the
	second paragraph of Section 4.3, the H notation includes the matches (Z), in notation
	and definition that seems to conflict with the statement just below Equation (10). This
	also comes into play in Equation (16)}

	\reply{%
	You are correct that the the comparison vectors are created (and therefore, patterns are assigned) without regard to $Z$. This notation was created purely to be able to denote when a record was left unlinked during the Gibbs Sampler as in Equation 16 in the submitted draft.

	I can see why this is confusing. Therefore, I have removed that line, and changed the notation for record $B_j$ being unmatched to just be''otherwise".
	}

\pointRaised{R}{%
	Third paragraph of Section 4.1: The authors claim they are computing “sufficient
	statistics”. What exactly are these statistics sufficient for?
	}

	\reply{%

We show in the revised Section 4.3 that $\tilde{\Gamma} = \{\mathcal{P}, \mathcal{R}, \mathcal{N} \}$ is a set of statistics that are sufficient for $m$, $u$, and $Z$.

{\color{red}{I drafted a whole bunch of material to argue more rigoursly that $\tilde{\Gamma} = \{\mathcal{P}, \mathcal{R}, \mathcal{N} \}$ is  set of sufficient statistics. If this argument seems weak, or is too much new material, I can also just keep what I had, and use less rigorous term "summary statistics" instead.}} \todo

}

\pointRaised{R}{%
First paragraph of Section 4.2: The authors state: “Posterior calculations still attribute
the appropriate weight to all records through the summary statistics…” What is meant
by the term “weight”? Which records are appropriately weighted – those in X2?}

\reply{%
	For each $j \in \{1, \ldots, n_B\}$, I use SEI to eliminate from memory the indices of  many records in $A$ that are unlikely matches for $B_j$ However, since I know the number of such records that exhibit each pattern $p$, I can still include their contribution to the likelihood through $N_{p_j}$.

	We have changed that sentence in the text to read ''All original record pairs still contribute to the likelihood through $\mathcal{N}$, and thus we can proceed with posterior inference through the memory reduced $\tilde{\Gamma}^{\text{SEI}} = \{\mathcal{P}, \mathcal{R}^{\text{SEI}}, \mathcal{N}\}$."

	This is clearer when looking at the conditional likelihoods in Equations 15 and 17 using $\tilde{\Gamma}$ in the revised Section 4.3. 
}

\pointRaised{R}{%
Second paragraph of Section 4.2: “and delete those comparison vectors”. Which are
“those” vectors?}

\reply{%
	We have revised that sentence to read ''We then conduct hashing, obtain the compressed $\tilde{\Gamma}^{ab}$ for later calculations, and delete the larger $\Gamma^{ab}$ from memory before continuing with the next chunk of data."
}

\pointRaised{R}{%
Where does $R^{SEI, cd}$ come into play in the partitioned algorithm presented in Equations
(13) and (14)? I recommend that the authors either refrain from suppressing the SEI
notation or further explain how the SEI algorithm has changed the quantities in these
equations.}

\reply{%
	The written expression in (13a) is correct. We then compress $r_{p_j}^{cd}$, and store in memory $r_{p_j}^{SEI, cd}$ Equation (14a) is conducted through $r_{p_j}^{SEI, cd}$. Equations (13b-c) and (14b-c) are exact, unchanged by SEI. We have clarified this.
}

\pointRaised{R}{%
Section 4.2: I appreciate the practical advice about choosing S for the SEI method. However,
this choice seems arbitrary in the absence of further discussion/evidence. Given that the
primary novelty of the manuscript is in methods to speed and otherwise improve computation, I
am surprised that this aspect of computational innovation is presented with virtually no
theoretical or empirical exploration. Presumably the SEI method has some sort of accuracy
trade-off, as the authors warn that linkage results may be “distorted” if S is low. However, this
trade-off is not quantified or even discussed in practical terms beyond the terse
recommendation to choose S=10.
}

\reply{%

{\color{red}{I'm not sure the best way to address this. Here is the where the situation arises:

Imagine file $B$ has one record $B_1$ for Brian Kundinger, and file $A$ has 20 nearly matching records for Brian Kundinger. Perhaps Brian A Kundinger, Brian B Kundinger, Brian C Kundinger, $\ldots$, ect. All of these records share the same agreement pattern (say $p = 2$) with $B_1$.

Without SEI, posterior probability would be divided evenly among all 20 near matches, such that the posterior probability would be at most $\frac{1}{20}$. If I chose $S=2$ however and I only retained 2 representatives in $r_{2_1}$, I could likely have one of those records having posterior probability above .5, and I would erroneously declare it a match.

So issues only arise in very degenerate situations. How do you think I should address this? Should I do a simulation showing degeneracy when S=1 or 2?
}
}
}
\pointRaised{R}{%
Page 5, 2 sentences before equation (4): I believe the sum should be of $I(Z_j \leq n_1)$, not $I(Z_j \leq n_1 +1)$.}

\reply{%
	That is correct. This has been corrected, and we apologize for the oversight. 
}

\pointRaised{R}{%
Equation (6a): The indices do not match the subscripts in the indicator function in each
summand, or their standard meaning in table 1.}

\reply{%
	{\color{red}{I do not understand this objection.}}
}

\pointRaised{R}{%
Generally, the authors seem to arbitrarily use upper and lower case z interchangeably in
function definitions.}

\reply{%
	{\color{red}{When writing, I attempted to use $Z$ when I was discussing a random quantity, and $z$ to reference a realized value. I haven't found anything that I see wrong. I think I'm going to need help with this.}}
}

\pointRaised{R}{%
Equations (8) and (9): Should the weights have superscript (s) (as the Zs do)?}

### 2023-05-02 ba_resubmit/kundinger_resubmit2.tex
response to reviewers

% Possible citation for bipartite matching. 
%\citep{fellegi_theory_1969, jaro1989, Winkler1988, belin_1995, larsen_2001, liseo_2011,  herzog2007data, gutman_bayesian_2013, sadinle_bayesian_2017}.

@@

Consider two data files $\bm{X}_1$ and $\bm{X}_2$ containing records $\{x_{1i}\}_{i=1}^{n_1}$ and $\{x_{2j}\}_{j=1}^{n_2}$ respectively. Without loss of generality, denote files such that $n_1 \geq n_2$. For ease of readability, we follow the convention established by \cite{sadinle_bayesian_2017} and say ``record $i \in \bm{X}_1$" rather than the more compact $x_{1i}$. Under bipartite matching, the set of matches across data files can be represented in two equivalent ways. First, we may use a matrix $\Delta \in \{0, 1\}^{n_1 \times n_2}$, where
\begin{align}
	\Delta_{ij} =
	\begin{cases}
		1, & \text{if records}\;  i \in \bm{X}_1 \; \text{and}\; j\in \bm{X}_2 \; \text{refer to the same entity}; \\
		0, & \text{otherwise}.\\
	\end{cases}
\end{align}
This sparse matrix representation can become cumbersome for large linkage tasks. More compactly, bipartite matching also can be viewed as a labeling $\bm{Z} = (Z_1, \ldots, Z_{n_2})$ for the records in $\bm{X}_2$ such that 
\begin{align}
	Z_{j} =
	\begin{cases}
		i, & \text{if records}\;  i \in \bm{X}_1 \; \text{and}\; j\in \bm{X}_2 \; \text{refer to the same entity}; \\
		n_1 + j, & \text{if record}\;  j \in \bm{X}_2 \; \text{does not have a match in}\; \bm{X}_1. \\
	\end{cases}
\end{align}
We can go back and forth between the two using $\Delta_{ij} = I(Z_j = i),$ where $I(\cdot) = 1$ when the expression inside the parentheses is true, and $I(\cdot) = 0$ otherwise.

Denote the set of matches by $\bm{M} = \{(i,j): i \in \bm{X}_1, j \in \bm{X}_2, \Delta_{ij} = 1\}$, and the set of non-matches by 
$\bm{U} =  \{(i,j): i \in \bm{X}_1, j \in \bm{X}_2, \Delta_{ij} = 0\}.$ The record linkage task can be viewed as identifying  $\bm{M}$ and  $\bm{U}.$ We refer to record pairs that are estimated as matches as ``links'' and to record pairs that are estimated as non-matches as ``non-links.''

@@

Intuitively, matching records (those that refer to the same entity) should be similar; records that are non-matching should be dissimilar. \cite{fellegi_theory_1969} proposed encoding this using a comparison vector $\bm{\gamma}_{ij}$ computed for each record pair $(i,j)$ in $\bm{X}_1 \times \bm{X}_2.$ Denote the number of criteria for comparing records by $F$, such that $\bm{\gamma}_{ij} = (\gamma_{ij}^1, \gamma_{ij}^2, \ldots, \gamma_{ij}^f, \ldots, \gamma_{ij}^F).$ We refer to set of all comparison vectors $\bm{\gamma}_{ij}$ as $\Gamma$.

@@

The simplest way to compare any particular feature for two records is to check for exact agreement, and this is commonly used for categorical features. For example, if zip code is linking feature $f$, we can set $\gamma_{ij}^f=1$ when the zip codes for record $i$ and $j$ agree exactly, and set $\gamma_{ij}^f=2$ when they do not. For numerical features, we can use absolute difference between the two feature values. For example, if age is linking field $f$, we can set $\gamma_{ij}^f = 1$ when the ages for record $i$ and $j$ match exactly, $\gamma_{ij}^f = 2$ when the ages for record $i$ and $j$ are within one year but not equal, and $\gamma_{ij}^f = 3$ when the ages are two or more years apart. For text features, such as names, we can use string distance metrics such as Levenstein or Jaro-Winkler distance \citep{cohen2003comparison}. We then set thresholds that allow us to represent comparisons through discrete levels of disagreement \citep{bilenko2006riddle, elmagarmid_duplicate_2007}.

More generally, let $\mathcal{S}_f(i,j)$ denote a similarity measure for feature $f$ of records $i$ and $j,$ where the range of $\mathcal{S}_f$ can be divided into $L_f$ intervals denoted by $I_{f1}, \ldots, I_{fL_f}$. Here, $I_{f1}$ represents the highest level of agreement (inclusive of complete agreement) and $I_{fL_f}$ represents the highest level of disagreement (including complete disagreement). Thus, we can construct comparison vectors such that $\gamma_{ij}^f = l \; \text{if} \; \mathcal{S}_f(i,j) \in I_{fl}.$ The choices of $I_{fl}$ are application specific, as we discuss in the simulation and case studies.

In the construction of comparison vectors, it is common to encounter missing information in record $i \in \bm{X}_1$ or $j \in \bm{X}_2$. As a result, the comparison vector $\mathbf{\gamma}_{ij}$ will have missing values. We assume that this missingness occurs at random (MAR, per \cite{LittleRubin2002}). With the MAR assumption, we can marginalize over the missing data, and do all computation simply using the observed data. To notate a missing value in any $\gamma_{ij}^f$, we use $I_{obs}(\gamma_{ij}^f)=1$ when $\gamma_{ij}^f$ is observed and $I_{obs}(\gamma_{ij}^f)=0$ otherwise.

@@

The seminal \cite{fellegi_theory_1969} model employs two independence assumptions: first, that comparison vectors are conditionally independent given their matching status, and second, that the matching status of the record pairs are independent. Using these assumptions, \cite{winkler_state_1999, jaro1989, larsen_2001, enamorado2019using} and others model the comparison data through mixture models of the form
\begin{subequations}
\begin{align}
	&\Gamma_{ij} = \gamma_{ij} \mid \Delta_{ij} = 1 \stackrel{iid}{\sim} \mathcal{M}(\bm{m}), \label{eqn:fs_model} \\
	&\Gamma_{ij} = \gamma_{ij} \mid \Delta_{ij} = 0  \stackrel{iid}{\sim} \mathcal{U}(\bm{u}), \label{eqn:m_dist}\\
	& \Delta_{ij}   \stackrel{iid}{\sim} \text{Bernoulli}(\lambda). \label{eqn:u_dist}
\end{align}
\end{subequations}
Here, $\mathcal{M}$ and $\mathcal{U}$ are the distributions for matching and non-matching record pairs, $\bm{m}$ and $\bm{u}$ are their respective sets of parameters, and $\lambda$ is the marginal probability that a record pair is a match. When using comparison vectors with discrete agreement levels, $\mathcal{M}$ and $\mathcal{U}$ are collections of independent multinomial distributions for each linkage feature. Accordingly, $\bm{m} = (\bm{m}_1, \ldots, \bm{m}_F)$, where $\bm{m}_f = (m_{f1}, \ldots, m_{fL_f})$ and $m_{fl} = P(\gamma_{ij}^f = l|\Delta_{ij} = 1)$ for all fields $f$ and agreement levels $l$. The $\bm{u}$ parameters are defined similarly, with $u_{fl} = P(\gamma_{ij}^f = l|\Delta_{ij} = 0)$.

%\textcolor{red}{Each record pair is then independently classified as a match or nonmatch according to the} likelihood ratio  
%\begin{align}
%	\label{eqn:wts}
%	w_{ij} = \frac{P(\gamma_{ij} \mid \Delta_{ij} = 1)}{P(\gamma_{ij} \mid \Delta_{ij} = 0)}.
%\end{align}
%This ratio is large when there is strong evidence of the pair being a match, and small otherwise. All record pairs with weights above a specified threshold are considered a match.

@@

%To address this issue, \cite{jaro1989} proposed solving the following linear sum assignment problem:
%\begin{align}
%	\label{eqn:jaro}
%	&\max_{\Delta} \sum_{i=1}^{n_1} \sum_{j=1}^{n_2} w_{ij} \Delta_{ij} 
%	\quad \text{subject to} \quad \Delta_{ij} \in \{0,1\}; \\
%	& \quad \quad \sum_{i=1}^{n_1}  \Delta_{ij}  \leq 1, j=1,2, \ldots n_2; \; \text{and} \notag\\
%	& \quad \quad\sum_{i=1}^{n_2}  \Delta_{ij}  \leq 1, i=1,2, \ldots n_1. \notag
%\end{align}
%
%The solution to this problem, for which several simple algorithms exist, is a bipartite matching that maximizes the sum of the Fellegi-Sunter weights among matched pairs. Jaro provided no theoretical justification for using this approach; however \cite{sadinle_bayesian_2017} recently showed that under certain conditions, (\ref{eqn:jaro}) is the maximum likelihood estimate for bipartite matching.

@@

The scale of linkage tasks possible through \texttt{BRL} is limited by the memory costs of storing $n_1 \times n_2$ comparison vectors for every pair of records across the two data files, and the speed of the linkage algorithm over those comparison vectors. One approach to reduce the number of comparisons is blocking, which places similar records into partitions, or ``blocks'' \citep{christen2019data}. In deterministic blocking, the modeler chooses fields thought to be highly reliable, and only compares records that agree on those fields. The record linkage method is  applied independently across all blocks, which can be done in parallel for additional speed gains. Of note, blocking on an unreliable field can lead to missed matches, making this form of blocking undesirable in some situations \citep{steorts_comparison_2014}.

>>>>

% Possible citation for bipartite matching. 
%\citep{fellegi_theory_1969, jaro1989, Winkler1988, belin_A995, larsen_2001, liseo_2011,  herzog2007data, gutman_bayesian_2013, sadinle_bayesian_2017}.

@@

%Consider two data files $A$ and $B$ containing records $\{x_{1i}\}_{i=1}^{n_A}$ and $\{x_{2j}\}_{j=1}^{n_B}$ respectively.

Consider two data files $A$ and $B$, consisting of records $A_i$ and $B_j$ respectively, where $i \in \{1, \ldots n_A\}$ and $j \in \{1, \ldots n_B\}$. Without loss of generality, denote files such that $n_A \geq n_B$. Under bipartite matching, the set of matches across data files can be represented in two equivalent ways. First, we may use a matrix $\Delta \in \{0, 1\}^{n_A \times n_B}$, where
\begin{align}
	\Delta_{ij} =
	\begin{cases}
		1, & \text{if records}\;  A_i \; \text{and}\; B_j \; \text{refer to the same entity}; \\
		0, & \text{otherwise}.\\
	\end{cases}
\end{align}
This sparse matrix representation can become cumbersome for large linkage tasks. More compactly, bipartite matching also can be viewed as a labeling $\bm{Z} = (Z_1, \ldots, Z_{n_B})$ for the records in $B$ such that 
\begin{align}
	Z_{j} =
	\begin{cases}
		i, & \text{if records}\;  A_i \; \text{and}\; B_j  \; \text{refer to the same entity}; \\
		n_A + j, & \text{if record}\;  B_j \; \text{does not have a match in}\; A. \\
	\end{cases}
\end{align}
We can go back and forth between the two using $\Delta_{ij} = I(Z_j = i),$ where $I(\cdot) = 1$ when the expression inside the parentheses is true, and $I(\cdot) = 0$ otherwise.

Denote the set of matches by $\bm{M} = \{(i,j): \Delta_{ij} = 1\}$, and the set of non-matches by 
$\bm{U} =  \{(i,j): \Delta_{ij} = 0\}.$ The record linkage task can be viewed as identifying  $\bm{M}$ and  $\bm{U}.$ We refer to record pairs that are estimated as matches as ``links'' and to record pairs that are estimated as non-matches as ``non-links.''

@@

Intuitively, matching records (those that refer to the same entity) should be similar; records that are non-matching should be dissimilar. \cite{fellegi_theory_1969} proposed encoding this using a comparison vector $\bm{\gamma}_{ij}$ computed for each record pair $(i,j)$ in $A \times B.$ Denote the number of criteria for comparing records by $F$, such that $\gamma_{ij} = (\gamma_{ij}^1, \ldots, \gamma_{ij}^F).$ We collect the vectors $\bm{\gamma}_{ij}$ as the comparison matrix $\Gamma \in \mathbb{R}^{n_A n_B \times F}$.

@@

The simplest way to compare any particular feature for two records is to check for exact agreement, and this is commonly used for categorical features. For example, if zip code is linking feature $f$, we can set $\gamma_{ij}^f=1$ when the zip codes for records $A_i$ and $B_j$ agree exactly, and set $\gamma_{ij}^f=2$ when they do not. For numerical features, we can use absolute difference between the two feature values. For example, if age is linking field $f$, we can set $\gamma_{ij}^f = 1$ when the ages for records $A_i$ and $B_j$ match exactly, $\gamma_{ij}^f = 2$ when the ages for records $A_i$ and $B_j$ are within one year but not equal, and $\gamma_{ij}^f = 3$ when the ages are two or more years apart. For text features, such as names, we can use string distance metrics such as Levenstein or Jaro-Winkler distance \citep{cohen2003comparison}. We then set thresholds that allow us to represent comparisons through discrete levels of disagreement \citep{bilenko2006riddle, elmagarmid_duplicate_2007}.

More generally, let $\mathcal{S}_f(i,j)$ denote a similarity measure for feature $f$ of records $A_i$ and $B_j,$ where the range of $\mathcal{S}_f$ can be divided into $L_f$ intervals denoted by $I_{f1}, \ldots, I_{fL_f}$. Here, $I_{f1}$ represents the highest level of agreement (including complete agreement) and $I_{fL_f}$ represents the highest level of disagreement (including complete disagreement). Thus, we can construct comparison vectors such that $\gamma_{ij}^f = l \; \text{if} \; \mathcal{S}_f(i,j) \in I_{fl}.$ The choices of $I_{fl}$ are application specific, as we discuss in the simulation and case studies.

{\color{red}{
In the construction of comparison vectors, it is common to encounter missing information in record $A_i$ or $B_j$. As a result, the comparison vector $\mathbf{\gamma}_{ij}$ will have missing values. We assume that this missingness occurs completely at random (MCAR, per \cite{LittleRubin2002}). With the MCAR assumption, we can marginalize over the missing data, and do all computation simply using the observed data. To notate a missing value in any $\gamma_{ij}^f$, we use $I_{obs}(\gamma_{ij}^f)=1$ when $\gamma_{ij}^f$ is observed and $I_{obs}(\gamma_{ij}^f)=0$ otherwise.}
}

@@

The seminal \cite{fellegi_theory_1969} model employs two independence assumptions: first, that comparison vectors are conditionally independent given their matching status, and second, that the matching status of the record pairs are independent. Using these assumptions, \cite{winkler_state_1999, jaro1989, larsen_2001, enamorado2019using} and others model the comparison data through mixture models of the form
\begin{subequations}
\begin{align}
	&\Gamma_{ij} = \gamma_{ij} \mid \Delta_{ij} = 1 \stackrel{iid}{\sim} \mathcal{M}(\bm{m}), \label{eqn:fs_model} \\
	&\Gamma_{ij} = \gamma_{ij} \mid \Delta_{ij} = 0  \stackrel{iid}{\sim} \mathcal{U}(\bm{u}), \label{eqn:m_dist}\\
	&\Delta_{ij}   \stackrel{iid}{\sim} \text{Bernoulli}(\lambda). \label{eqn:u_dist}
\end{align}
\end{subequations}
Here, $\mathcal{M}$ and $\mathcal{U}$ are the distributions for matching and non-matching record pairs, $\bm{m}$ and $\bm{u}$ are their respective sets of parameters, and $\lambda$ is the marginal probability that a record pair is a match. When using comparison vectors with discrete agreement levels, $\mathcal{M}$ and $\mathcal{U}$ are collections of independent multinomial distributions for each linkage feature. Accordingly, $\bm{m} = (\bm{m}_1, \ldots, \bm{m}_F)$, where $\bm{m}_f = (m_{f1}, \ldots, m_{fL_f})$ and $m_{fl} = P(\gamma_{ij}^f = l|\Delta_{ij} = 1)$ for all fields $f$ and agreement levels $l$. The $\bm{u}$ parameters are defined similarly, with $u_{fl} = P(\gamma_{ij}^f = l|\Delta_{ij} = 0)$.

@@

%To address this issue, \cite{jaro1989} proposed solving the following linear sum assignment problem:
%\begin{align}
%	\label{eqn:jaro}
%	&\max_{\Delta} \sum_{i=1}^{n_A} \sum_{j=1}^{n_B} w_{ij} \Delta_{ij} 
%	\quad \text{subject to} \quad \Delta_{ij} \in \{0,1\}; \\
%	& \quad \quad \sum_{i=1}^{n_A}  \Delta_{ij}  \leq 1, j=1,2, \ldots n_B; \; \text{and} \notag\\
%	& \quad \quad\sum_{i=1}^{n_B}  \Delta_{ij}  \leq 1, i=1,2, \ldots n_A. \notag
%\end{align}
%
%The solution to this problem, for which several simple algorithms exist, is a bipartite matching that maximizes the sum of the Fellegi-Sunter weights among matched pairs. Jaro provided no theoretical justification for using this approach; however \cite{sadinle_bayesian_2017} recently showed that under certain conditions, (\ref{eqn:jaro}) is the maximum likelihood estimate for bipartite matching.

@@

The scale of linkage tasks possible through \texttt{BRL} is limited by the memory costs of storing $n_A \times n_B$ comparison vectors for every pair of records across the two data files, and the speed of the linkage algorithm over those comparison vectors. One approach to reduce the number of comparisons is blocking, which places similar records into partitions, or ``blocks'' \citep{christen2019data}. In deterministic blocking, the modeler chooses fields thought to be highly reliable, and only compares records that agree on those fields. The record linkage method is  applied independently across all blocks, which can be done in parallel for additional speed gains. Of note, blocking on an unreliable field can lead to missed matches, making this form of blocking undesirable in some situations \citep{steorts_comparison_2014}.
